Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries

### 2018-12-31 docs/manuscript.tex
Made last changes and re-submitted

\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in small-scale fisheries}

@@

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Small-scale fisheries face
great challenges since they are difficult to monitor, enforce, and
manage. Combining territorial user rights for fisheries (TURF) with
no-take marine reserves to create TURF-reserves can improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last 12
years, 18 old and new community-based Mexican TURF-reserves gained legal
recognition thanks to a 2014 regulation; their effectiveness has not
been formally evaluated. We combine causal inference techniques and the
Social-Ecological Systems framework to provide a holistic evaluation of
community-based TURF-reserves in three coastal communities in Mexico. We
find that while reserves have not yet achieved their stated goal of
increasing the density of lobster and other benthic invertebrates, they
continue to receive support from the fishing communities. A lack of
clear ecological and socioeconomic effects likely results from a
combination of factors. First, some of these reserves might be too young
for the effects to show. Second, the reserves are not large enough to
protect mobile species, like lobster. Third, variable and extreme
oceanographic conditions have impacted harvested populations. Fourth,
local fisheries are already well managed, and it is unlikely that
reserves might have a detectable effect in catches. However, these
reserves may provide a foundation for establishing additional, larger
marine reserves needed to effectively conserve mobile species.

@@

TURFs are a fisheries management tool in which a well defined group of
fishers have exclusive access to an explicitly delimited portion of the
ocean. They promote a sense of stewardship and incentivise resource
users to sustainably manage their resources
\citep{gelcich_2008,costello_2010,mccay_2014}. On the other hand,
no-take marine reserves (marine reserves from hereinafter) are areas
where all extractive activities are off-limits. These can be implemented
to protect biodiversity but also as fishery management tools that
restrict fishing effort and gears and therefore aid in the recovery of
marine stocks. These instruments can be combined by establishing a
marine reserve within a TURF, thus making them TURF-reserves
\citep{afflerbach_2014,gelcich_2015,lester_2017}.

Conservation science has shown how marine reserves lead to increased
biomass, species richness, and abundance within the protected regions
\citep{lester_2009}, and that these may have a series of additional
benefits like climate change mitigation, protection from environmental
variability, and fisheries benefits
\citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
research on TURFs has shown that these areas have higher abundance of
targeted species than sites operating under open access and even similar
to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The
benefits resulting from reserves established within TURFs (\emph{i.e.}
TURF-reserves) should be captured exclusively by the group of fishers
with exclusive access \citep{gelcich_2015}. Although in theory these
systems are successful \citep{smallhornwest_2018}, there is little
empirical evidence of their effectiveness and the drivers of their
success.

TURF-reserve systems are inherently intricate social-ecological systems,
and their effectiveness must depend on how environmental and social
factors combine and interact \citep{gelcich_2015}. This is especially
important in social-ecological coastal systems dominated by close
interactions and feedbacks between people and natural resources
\citep{ostrom_2009}. There is a growing body of literature focusing on
the relations between socioeconomic and governance structures and
reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}.
However, to our knowledge, no studies exist that evaluate TURF-reserves
from both a social and ecological perspective.

Recent norms in fisheries regulation in Mexico provide a ripe
opportunity to study the effectiveness of community-based TURF-reserves
in small-scale fisheries. In Mexico, a norm created in 2014 allows
fishers to request legal recognition of community-based reserves as
``Fish Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since
2012, old and new marine reserves have gained legal recognition as
Fishing Refuges. Of these, 18 were originally implemented as within
TURFs. However, their effectiveness has not yet been formally evaluated
and reported in the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological
Systems (SES) framework to evaluate community-based TURF-reserves in
three coastal communities in Mexico. These three case studies span a
range of ecological and social conditions representative of different
regions of Mexico. The objective of this work is twofold. First, to
provide a holistic evaluation of the effectiveness of community-based
TURF-reserves in terms of the changes in biological and socioeconomic
indicators and the governance settings under which these develop, which
may inform similar processes in other countries. Second, to identify
opportunities where improvement or adjustment might lead to increased
effectiveness. We draw from lessons learned in these three case studies
and provide management recommendations to maximize the effectiveness of
community-based TURF-reserves in small-scale fisheries where this tool
is used to manage and rebuild their coastal fisheries.

@@

In an effort to bridge this normative gap, Mexican Civil Society
Organizations (CSOs) served as a link between fishers and government,
and created a legal framework that solves this governance issue
(\emph{i.e.} Fish Refuges; \citet{nom}). Fish Refuges can be implemented
as temporal or partial reserves, which can protect one, some, or all
resources within their boundaries. One of the ways in which fishing
communities have taken advantage of this new tool is by implementing
temporal marine reserves within their TURFs with a defined expiration
date (often 5 years). After these five years have passed, fishers have
the opportunity to open the reserves to fishing or continue to have
them. Our work focuses on Fish Refuges implemented as community-based
TURF-reserves that occur in small-scale fisheries.

The common setup of community-based TURF-reserves in Mexico is the
following. Fishers from a given community are assembled in fishing
cooperatives which have exclusive fishing rights over a spatially
delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
\ref{fig:map}A). Each TURF is elusively fished by one cooperative, and
each community usually hosts no more than one cooperative. The profits
from each TURF are shared amongst all fishers from the cooperative.
Fishing cooperatives interested in implementing marine reserves work
with CSOs to implement marine reserves within their TURFs (\emph{i.e.}
TURF-reserves). Fishers then ask the government to grant legal
recognition to their TURF-reserves as Fish Refuges following a series of
studies outlined in the regulation \citep{nom}.

@@

We evaluate three community-based no-take TURF-reserves implemented in
Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig
\ref{fig:map}A). The first one was created by the \emph{Buzos y
Pescadores de la Baja California} fishing cooperative, located in Isla
Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The
main fishery in the island is the spiny lobster (\emph{Panulirus
interruptus}), but other resources like finfish, sea cucumber, read sea
urchin, snail, and abalone are also an important source of income. In
2006, the community decided to implement two marine reserves within
their fishing grounds to protect commercially important invertebrate
species; mainly lobster and abalone. These reserves obtained legal
recognition in 2018 \citep{dof_website_2018}.

The other two TURF-reserves are located in Maria Elena and Punta
Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
Isla Nativdad, which hosts a well established fishing community, Maria
Elena is a fishing camp --visited intermittently during the fishing
season-- belonging to the \emph{Cozumel} fishing cooperative; Punta
Herrero is home to the \emph{José María Azcorra} fishing cooperative,
and similar to Isla Natividad hosts a local community. Their main
fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but
they also target finfish in the off-season. Maria Elena and Punta
Herrero established eight and four marine reserves in 2012 and 2013,
respectively. These reserves have been legally recognized as Fishing
Refuges since their original implementation
\citep{dof_website_2012,dof_website_2013} and re-enactments
\citep{dof_website_2017b}.

@@

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control sites. Reserve sites are those within the reserves, and thus
no fishing takes place. Control sites are areas that meet the following
criteria: i) habitat characteristics are similar to the corresponding
reserves, ii) presumably had a similar probability of being selected as
reserves during the design phase, iii) are located within the TURF,
where fishing occurs, and iv) Are not directly adjacent to the reserves.
We focus our evaluation on sites where data are available for reserve
and control sites, before and after the implementation of the reserve.
This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
sampling design that allows us to capture and control for temporal and
spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally
attribute the changes to the reserve
\citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.

@@

Socioeconomic data come from landing receipts reported to the National
Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
(Kg) and revenues (MXP) for TURF-managed cooperatives with and without
marine reserves. In this case our treated unit are the cooperatives
(\emph{i.e.} communities) that have implemented a reserve within their
TURF, and the controls are nearby communities that have a TURF but did
not implement a reserve. Cooperatives incorporated in this analysis
belong to larger regional-level Cooperative Federations, and are exposed
to the same markets and institutional frameworks, making them plausible
controls \citep{mccay_2017,ayer_2018}. Landings and revenues were
aggregated at the cooperative-year level, and revenues were adjusted to
represent 2014 values by the Consumer Price Index for Mexico
\citep{oecd_2017}. A table with summary statistics for this data is
provided in the supplementary materials \textbf{(Table S3, Figure S5)}.

Data for the evaluation of the SES were collected at the community-level
from official documents used in the creation and designation of the
marine reserves. These includes the technical studies that the
cooperatives submit when they request recognition of their reserves, as
well as the official enactments
\citep{dof_website_2012,dof_website_2013,dof_website_2018}. We also
complimented information based on the authors' experience and knowledge
of the communities. We collected information on the Resource Systems,
Resource Units, Actors, and Governance System (Table \ref{table:ses}).
The next sections further describe the variables and indicators chosen
for each of these.

@@

\begin{equation}
I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t,j}
\label{eqn:reg_bio}
\end{equation}

Where year-level fixed effects capturing a temporal trend are
represented by \(\gamma_t Year_t\), and \(\beta Zone_i\) captures the
difference between reserve (\(Zone = 1\)) and control (\(Zone = 0\))
sites. The effect of the reserve is captured by the \(\lambda_t\)
coefficient, and represents the difference observed between the control
site before the implementation of the reserve and the treated sites at
time \(t\) after controlling for other time and space variations
(\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Therefore, we
would expect this term to be positive if the indicator increases because
of the reserve implementation. Finally, \(\epsilon_{i,t,j}\) represents
the error term of the regression.

Socioeconomic indicators are evaluated with a similar approach. Due to
data constrains, we only evaluate socioeconomic data for Isla Natividad
(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
used as counterfactuals that allow us to control for unobserved
time-invariants. Each focal community (Isla Natividad and Maria Elena)
has three counterfactual communities.

\begin{equation}
I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t,j}
\label{eqn:soc_reg}
\end{equation}

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio},
but in this case the \(Treated\) dummy variable indicates if the
community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)).
These regression models allow us to establish a causal link between the
implementation of marine reserves and the observed trends by accounting
for temporal and site-specific dynamics \citep{depalma_2018}. All model
coefficients were estimated via ordinary least-squares and
heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
analyses were performed in R version 3.5.1 and R Studio version 1.1.456
\citep{R_2018}. All data and code needed to reproduce our analyses are
available in a \href{https://github.com/jcvdav/ReserveEffect}{GitHub
repository}.

@@

While many ecology studies have used BACI sampling designs and
respective analyses (\emph{e.g.} \citet{stewartoaten_1986}), few
conservation studies have done so to evaluate the effect of an
intervention (\emph{e.g.}
\citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted
in a call for more robust analyses in conservation science
\citep{guidetti_2002,ferraro_2006}. Our approach to evaluate the
temporal and spatial changes provides a more robust measure of reserve
effectiveness. For example, we capture previously described patterns
like the rapid increase observed for lobster densities in Isla Natividad
on the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), a year
after the hypoxia events described by \citet{micheli_2012-EU}, which
caused mass mortality of sedentary organisms such as abalone and sea
urchins, but not lobster and finfish. Yet, our empirical approach
assumes control sites are a plausible counterfactual for treated sites.
This implies that treated sites would have followed the same trend as
control sites, had the reserves not been implemented. Nonetheless,
temporal trends for each site don't show any significant increases
(Figures S1-S4), supporting our findings of lack of change in the
indicators used.

Our analyses of socioeconomic indicators has two limitations. First, we
only look at landings and revenues by landings for communities with and
without TURF-reserves. There are a number of other possible indicators
that could show a change due to the implementation of the reserve.
Notably, one often cited in the literature is additional benefits, such
as tourism \citep{viana_2017}. However, it is unlikely that the
evaluated communities will experience tourism benefits due to their
remoteness and the lack of proper infrastructure to sustain tourism. A
second limitation of our socioeconomic analysis is that we do not
observe effort data, which may mask the effect of the reserve. For
example, if catches remain relatively unchanged but fishing effort
decreased, that would imply a larger catch per unit effort and higher
profitability.

@@

Even if reserves had appropriate sizes and were placed in optimal
locations, there are other plausible explanations for the observed
patterns. For instance, marine reserves are only likely to provide
fisheries benefits if initial population sizes are low and the fishery
is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
fisheries were certified by the Marine Stewardship Council and lobster
fisheries are managed via species-specific minimum catch sizes, seasonal
closures, protection of ``berried'' females, and escapement windows
where traps are allowed \citep{dof_website_1993}. It is uncertain
whether such a well-managed fishery will experience additional benefits
from marine reserves; reserves implemented in TURFs where fishing
pressure is already optimally managed will still show a trade-off
between fisheries and conservation objectives \citep{lester_2017}.
Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have
greater biomass and richness than areas operating under open access.
This might reduce the difference between indicators from the TURF and
reserve sites, making it difficult to detect such a small change.
Further research should focus on evaluating sites in the reserve, TURF,
and open access areas or similar Fish Refuges established without the
presence of TURFs where the impact of the reserves might be larger.

@@

While the evaluated reserves have failed to provide fishery benefits up
to now, there are a number of additional ecological, fisheries, and
social benefits. Marine reserves provide protection to a wider range of
species and vulnerable habitat. Previous research focusing on these
specific sites has shown that they serve as an insurance mechanism
against uncertainty and errors in fisheries management, as well as mild
environmental shocks
\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
Self-regulation of fishing effort can serve as a way to compensate for
future declines associated to environmental variation
\citep{finkbeiner_2018}. Furthermore, embarking in a marine conservation
project can bring the community together, which promotes social cohesion
and builds social capital as reported by \citet{fulton_2019}. Showing
commitment to marine conservation and sustainable fishing practices has
allowed fishers to have greater bargaining power and leverage over
fisheries management \citep{prezramrez_2012}. These additional benefits
might explain why communities show a positive perception about their
performance and continue to support their presence by re-implementing
the reserves \citep{ayer_2018}.

In terms of the fisheries regulation in Mexico, our work only evaluates
Fish Refuges established within TURFs. Future research should aim at
evaluating other Fish Refuges that have also been established as
bottom-up processes but without the presence of TURFs (\emph{e.g.}
\citet{dof_websiteC_2012}), others established through top-down
processes (\emph{i.e.} \citet{dof_websiteU_2018}), as well as the
relationship between governance and effectiveness across this gradient
of approaches.

Community-based TURF-reserves in small-scale fisheries may be helpful
conservation and fishery management tools when appropriately implemented
\citep{gelcich_2015}. We must promote bottom-up design and
implementation processes like the ones in the evaluated reserves, but
without setting design principles aside. Having full community support
surely represents an advantage, but it is important that community-based
TURF-reserves meet essential design principles such as size and
placement so as to maximize their effectiveness. For the particular case
of the ones we evaluate, the possibility of expanding reserves or
merging existing polygons into larger areas should be evaluated and
proposed to the communities.

@@

The authors wish to acknowledge Imelda Amador for contributions on the
governance data, as well as pre-processing biological data. This study
would have not been possible without the effort by members of the
fishing communities here mentioned, who participated in the
data-collection process.

>>>>

\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in Mexican small-scale
fisheries}

@@

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Small-scale fisheries face
great challenges since they are difficult to monitor, enforce, and
manage. Combining territorial use rights for fisheries (TURF) with
no-take marine reserves to create TURF-reserves can improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last 12
years, 18 old and new community-based Mexican TURF-reserves gained legal
recognition thanks to a regulation passed in 2014; their effectiveness
has not been formally evaluated. We combine causal inference techniques
and the Social-Ecological Systems framework to provide a holistic
evaluation of community-based TURF-reserves in three coastal communities
in Mexico. We find that while reserves have not yet achieved their
stated goal of increasing the density of lobster and other benthic
invertebrates, they continue to receive support from the fishing
communities. A lack of clear ecological and socioeconomic effects likely
results from a combination of factors. First, some of these reserves
might be too young for the effects to show. Second, the reserves are not
large enough to protect mobile species, like lobster. Third, variable
and extreme oceanographic conditions have impacted harvested
populations. Fourth, local fisheries are already well managed, and it is
unlikely that reserves might have a detectable effect in catches.
However, these reserves may provide a foundation for establishing
additional, larger marine reserves needed to effectively conserve mobile
species.

@@

TURFs are a fisheries management tool in which a well defined group of
fishers (\emph{e.g.} fishing cooperatives) have exclusive access to an
explicitly delimited portion of the ocean. They promote a sense of
stewardship and incentivise resource users to sustainably manage their
resources \citep{gelcich_2008,costello_2010,mccay_2014}. On the other
hand, no-take marine reserves (marine reserves from hereinafter) are
areas where all extractive activities are off-limits. These can be
implemented to protect biodiversity but also as fishery management tools
to aid in the recovery of marine stocks. These instruments can be
combined by establishing a marine reserve within a TURF, thus making
them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.

Conservation science has shown how marine reserves lead to increased
biomass, species richness, and abundance within the protected regions
\citep{lester_2009}, and that these may have a series of additional
benefits like climate change mitigation, protection from environmental
variability, and fisheries benefits
\citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
research on TURFs has shown that these areas have higher abundance of
targeted species than sites operating under open access and even similar
to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The
benefits resulting from reserves established within TURFs (\emph{i.e.}
TURF-reserves) should be captured exclusively by the group of fishers
with exclusive access \citep{gelcich_2015}. Although in theory these
systems are successful \citep{smallhornwest_2018}, there is little
empirical evidence of their effectiveness and the drivers of their
success. Moreover, TURF-reserve systems are inherently intricate
social-ecological systems, and their effectiveness must depend on how
environmental and social factors combine and interact
\citep{ostrom_2009,gelcich_2015}. It is therefore important to consider
not only the indicators of interest, bu also the governance settings
under which the reserves occur.

Recent norms in fisheries regulation in Mexico provide a ripe
opportunity to study the effectiveness of community-based TURF-reserves
in small-scale fisheries. In Mexico, a norm created in 2014 allows
fishers to request legal recognition of community-based reserves as
``Fish Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since
2012, old and new marine reserves have gained legal recognition as Fish
Refuges. Of these, 18 were originally implemented within TURFs. However,
their effectiveness has not yet been formally evaluated and reported in
the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological
Systems (SES) framework to evaluate community-based TURF-reserves in
three coastal communities in Mexico. The objective of this work is
twofold. First, to provide a holistic evaluation of the effectiveness of
community-based TURF-reserves in terms of the changes in biological and
socioeconomic indicators and the governance settings under which these
develop, which may inform similar processes in other countries. Second,
to identify opportunities where improvement or adjustment might lead to
increased effectiveness. We draw from lessons learned in these three
case studies and provide management recommendations to maximize the
effectiveness of community-based TURF-reserves in small-scale fisheries
where this tool is used to manage and rebuild their coastal fisheries.

@@

In an effort to bridge this normative gap, Mexican Civil Society
Organizations (CSOs) served as a link between fishers and government,
and created a legal framework that solves this governance issue: Fish
Refuges \citep{nom}. Fish Refuges can be implemented as temporal or
partial reserves, which can protect one, some, or all resources within
their boundaries. One of the ways in which fishing communities have
taken advantage of this new tool is by implementing temporal marine
reserves within their TURFs with a defined expiration date (often 5
years). When the expiration date is reached, fishers can chose to open
the reserves to fishing or re-establish them. Our work focuses on Fish
Refuges implemented as community-based TURF-reserves that occur in
small-scale fisheries.

The common setup of community-based TURF-reserves in Mexico is the
following. Fishers from a given community are assembled in fishing
cooperatives which have exclusive fishing rights over a spatially
delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
\ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and
each community usually hosts no more than one cooperative. The profits
from each TURF are shared amongst all fishers from the cooperative.
Fishing cooperatives interested in implementing marine reserves work
with CSOs to implement marine reserves within their TURFs (\emph{i.e.}
TURF-reserves). Fishers then ask the government to grant legal
recognition to their TURF-reserves as Fish Refuges following a series of
studies outlined in the regulation \citep{nom}.

@@

We evaluate three community-based no-take TURF-reserves implemented in
Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig
\ref{fig:map}A). The first one was created by the \emph{Buzos y
Pescadores de la Baja California} fishing cooperative, located in Isla
Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The
main fishery in the island is the spiny lobster (\emph{Panulirus
interruptus}), but other resources like finfish, sea cucumber, read sea
urchin, snail, and abalone are also an important source of income. In
2006, the community decided to implement two marine reserves within
their fishing grounds. The objective of these reserves was ``to protect
commercially important invertebrate species''; mainly lobster and
abalone. These reserves obtained legal recognition in 2018
\citep{dof_website_2018}.

The other two TURF-reserves are located in Maria Elena and Punta
Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
Isla Nativdad, which hosts a well established fishing community, Maria
Elena is a fishing camp --visited intermittently during the fishing
season-- belonging to the \emph{Cozumel} fishing cooperative; Punta
Herrero is home to the \emph{José María Azcorra} fishing cooperative,
and similar to Isla Natividad hosts a local community. Their main
fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but
they also target finfish in the off-season. Maria Elena and Punta
Herrero established eight and four marine reserves in 2012 and 2013,
respectively. These reserves have been legally recognized as Fishing
Refuges since their original implementation
\citep{dof_website_2012,dof_website_2013} and subsequent
re-establishments \citep{dof_website_2017b}.

@@

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control sites. Reserve sites are areas where no fishing occurs.
Control sites are areas that meet the following criteria: i) habitat
characteristics are similar to the corresponding reserves, ii)
presumably had a similar probability of being selected as reserves
during the design phase, iii) are located within the TURF, where fishing
occurs, and iv) Are not directly adjacent to the reserves. We focus our
evaluation on sites where data are available for reserve and control
sites, before and after the implementation of the reserve. This provides
us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design
that allows us to capture and control for temporal and spatial dynamics
\citep{stewartoaten_1986,depalma_2018} and causally attribute the
changes to the reserve
\citep{francinifilho_2008,Villasenor-Derbez_2018}.

@@

Socioeconomic data come from landing receipts reported to the National
Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
(Kg) and revenues (MXP) for TURF-managed cooperatives with and without
marine reserves. In this case our treated unit are the cooperatives
(\emph{i.e.} communities) that have implemented a reserve within their
TURF, and the controls are nearby communities that have a TURF but did
not implement a reserve. Cooperatives incorporated in this analysis
belong to larger regional-level Cooperative Federations, and are exposed
to the same markets and institutional frameworks, making them plausible
controls \citep{mccay_2017,ayer_2018}. Landings and revenues were
aggregated at the cooperative-year level, and revenues were adjusted to
represent 2014 values by the Consumer Price Index for Mexico
\citep{oecd_2017}. A table with summary statistics for this data is
provided in the supplementary materials (Table S3, Figure S5).

Data for the evaluation of the SES were collected at the community-level
from official documents used in the design, creation, and designation of
the marine reserves. These include the technical studies that the
cooperatives submit when they request recognition of their reserves, as
well as the official enactments
\citep{dof_website_2012,dof_website_2013,dof_website_2018}. We also
complimented information based on the authors' experience and knowledge
of the communities. We collected information on the Resource Systems,
Resource Units, Actors, and Governance System (Table \ref{table:ses}).
The next section further describe the variables and indicators chosen
for each of these.

@@

\begin{equation}
I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t}
\label{eqn:reg_bio}
\end{equation}

Where year-level fixed effects capturing a temporal trend are
represented by \(\gamma_t Year_t\), and \(\beta Zone_i\) captures the
difference between reserve (\(Zone = 1\)) and control (\(Zone = 0\))
sites. The effect of the reserve is captured by the \(\lambda_t\)
coefficient, and represents the difference observed between the control
site before the implementation of the reserve and the treated sites at
time \(t\) after controlling for other time and space variations
(\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Therefore, we
would expect this term to be positive if the indicator increases because
of the reserve. Finally, \(\epsilon_{i,t}\) represents the error term of
the regression.

Socioeconomic indicators are evaluated with a similar approach. Due to
data constrains, we only evaluate socioeconomic data for Isla Natividad
(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
used as counterfactuals that allow us to control for unobserved
time-invariants. Each focal community (\emph{i.e.} Isla Natividad and
Maria Elena) has three counterfactual communities.

\begin{equation}
I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t}
\label{eqn:soc_reg}
\end{equation}

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio},
but in this case the \(Treated\) dummy variable indicates if the
community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)).
These regression models allow us to establish a causal link between the
implementation of marine reserves and the observed trends by accounting
for temporal and site-specific dynamics \citep{depalma_2018}. We fit one
model per community and indicators (\emph{e.g.} there are three models
for lobster density, one for each community) for a total of 12
biological model fits and four socioeconomic model fits. Model
coefficients were estimated via ordinary least-squares and used
heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
analyses were performed in R version 3.5.1 and R Studio version 1.1.456
\citep{R_2018}. All data and code needed to reproduce our analyses are
available in a GitHub repository at:
\url{https://github.com/jcvdav/ReserveEffect}.

@@

While many ecology studies have used BACI sampling designs and
respective analyses (\emph{e.g.} \citet{stewartoaten_1986}), few
conservation studies have done so to evaluate the effect of an
intervention (\emph{e.g.}
\citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted
in a call for more robust analyses in conservation science
\citep{guidetti_2002,ferraro_2006}. Our approach to evaluate the
temporal and spatial changes provides a more robust measure of reserve
effectiveness, and captures previously described patterns. For example,
the rapid increase observed for lobster densities in Isla Natividad on
the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), occurs a
year after the hypoxia events described by \citet{micheli_2012-EU},
which caused mass mortality of sedentary organisms such as abalone and
sea urchins, but not lobster and finfish. The use of causal inference
techniques may help us support evidence-based conservation.

Our analyses of socioeconomic indicators has two limitations. First, we
only look at landings and revenues by landings for communities with and
without TURF-reserves. There are a number of other possible indicators
that could show a change due to the implementation of the reserve.
Notably, one often cited in the literature is additional benefits, such
as tourism \citep{viana_2017}. However, it is unlikely that the
evaluated communities will experience tourism benefits due to their
remoteness and the lack of proper infrastructure to sustain tourism. A
second limitation of our socioeconomic analysis is that we do not
observe effort data, which may mask the effect of the reserve. For
example, if catches remain relatively unchanged but fishing effort
decreased, that would imply a larger catch per unit effort and thus
higher profitability.

@@

Even if reserves had appropriate sizes and were placed in optimal
locations, there are other plausible explanations for the observed
patterns. For instance, marine reserves are only likely to provide
fisheries benefits if initial population sizes are low and the fishery
is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
fisheries were certified by the Marine Stewardship Council and are
managed via species-specific minimum catch sizes, seasonal closures,
protection of ``berried'' females, and escapement windows where traps
are allowed \citep{dof_website_1993}. It is uncertain whether such a
well-managed fishery will experience additional benefits from marine
reserves; reserves implemented in TURFs where fishing pressure is
already optimally managed will still show a trade-off between fisheries
and conservation objectives \citep{lester_2017}. Furthermore,
\citet{gelcich_2008} have shown that TURFs alone can have greater
biomass and richness than areas operating under open access. This might
reduce the difference between indicators from the TURF and reserve
sites, making it difficult to detect such a small change. Further
research should focus on evaluating sites in the reserve, TURF, and open
access areas or similar Fish Refuges established without the presence of
TURFs where the impact of the reserves might be greater.

@@

While the evaluated reserves have failed to provide fishery benefits up
to now, there are a number of additional ecological, fisheries, and
social benefits. Marine reserves provide protection to a wider range of
species and vulnerable habitat. Previous research focusing on these
specific sites has shown that they serve as an insurance mechanism
against uncertainty and errors in fisheries management, as well as mild
environmental shocks
\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
Self-regulation of fishing effort can serve as a way to compensate for
future declines associated to environmental variation
\citep{finkbeiner_2018}. Furthermore, embarking in a marine conservation
project can bring the community together, which promotes social cohesion
and builds social capital \citep{fulton_2019}. Showing commitment to
marine conservation and sustainable fishing practices has allowed
fishers to have greater bargaining power and leverage over fisheries
management \citep{prezramrez_2012}. These additional benefits might
explain why communities show a positive perception about their
performance and continue to support their presence by re-implementing
the reserves \citep{ayer_2018}.

Community-based TURF-reserves in small-scale fisheries may be helpful
conservation and fishery management tools when appropriately implemented
\citep{gelcich_2015}. We must promote bottom-up design and
implementation processes like the ones in the evaluated reserves, but
without setting design principles aside. Having full community support
surely represents an advantage, but it is important that community-based
TURF-reserves meet essential design principles such as size and
placement so as to maximize their effectiveness. Furthermore, consvation
and advocacy groups should consider the opportunity costs of such
interventions (\emph{sensu} \citet{smith_2010}) and evaluate the
potential of other approaches that may yield similar benefits.

In terms of the fisheries regulation in Mexico, our work only evaluates
Fish Refuges established within TURFs. Future research should aim at
evaluating other Fish Refuges established as bottom-up processes but
without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
others established through top-down processes (\emph{i.e.}
\citet{dof_websiteU_2018}), as well as the relationship between
governance and effectiveness across this gradient of approaches. For the
particular case of the reserves that we evaluate, the possibility of
expanding reserves or merging existing polygons into larger areas should
be evaluated and proposed to the communities.

@@

The authors wish to acknowledge Imelda Amador for contributions on the
governance data, as well as pre-processing biological data. This study
would have not been possible without the effort by members of the
fishing communities here mentioned, who participated in the
data-collection process. The authors wish to acknowledge comments by the
reviewrs and editor, which significantly improved the quality of this
work.
